Stream of Consciousness Literature: A 'Joyceless' Linguistic Landscape

### 2024-04-05 paper/paper.qmd
paper update

---
title: "Stream of Consciousness Literature: A 'Joyceless' Linguistic Landscape" 
subtitle: "Exploring Word Frequency, Sentiment Value and Mental Health Themes in the Works of Joyce, Woolf, Proust, Mansfield and Eliot from Project Gutenberg"
author: Quang Mai
thanks: "Code and data are available at: https://github.com/ponolite/stream-consciousness-language.git"
date: today
date-format: long
abstract: "This project focuses on understanding the language used by renowned stream of consciousness (SOC) authors James Joyce, Virginia Woolf, Marcel Proust, Katherine Mansfield and T.S Eliot. By conducting word frequency analysis and sentiment analysis on these authors' nine novels, this paper aims to uncover shared linguistic patterns and gain insights into the authors' mental states. Ultimately, this paper attempts to offer insights into themes of self-identity, anxiety, disassociation and existential contemplation within Western society and its literary circle from late 19th to mid-20th century. (add one sentence on main results)"
format: pdf
number-sections: true
bibliography: references.bib
toc: true
---

@@

* **Word Frequency:** This variable captures the repetition of a single word (unigram), two-words combination (bigram) or three-words combination (trigram) throughout a SOC novel text, providing us with a thematic understanding of SOC literature.
* **Sentiment Value:** This variable enables us to analyze how every word is usually perceived emotionally, whether it be positively or negatively. In additon,    in childlessness over time, encompassing the chosen time frame for the survey.

Out of two variables used, the first one, 'Word Frequency' usually captured as `n` or `frequency` in datasets, is directly quantified through tokenizing the novel texts using the packages @rTidytext [@youtube1] [@youtube2]. To do this, I first downloaded all nine novel texts from @gutenberg, and leveraged functions such as `unnest_tokens()` from @rTidytext to mine the texts, or separating it into individual words. Finally, I used `count()` to quantify the word frequency.

The second variable used, 'Sentiment Value', is based on three English-based, general-purpose  "word-emotion and word-polarity association lexicons", sourced from @sentiment's expansive research along with efforts from Finn Årup Nielsen and Bing Liu and collaborators [@rTidytext].

The three general-purpose lexicon that contributes to my sentiment analysis are [@rTidytext]:

@@

In terms of measuring 'Sentiment Value', all three general-purpose lexicons are compiled through crowd-sourcing and directly surveying the public on how each word is emotionally perceived

A survey sample of how the word 'startle' is compiled within the 'nrc' lexicon is presented below [@sentiment]:

>>>>

---
title: "Stream of Consciousness Literature: A 'Joyceless' Linguistic Landscape" 
subtitle: "Exploring Word Frequency, Sentiment Value and Mental Health Themes in the Works of Joyce, Woolf, Proust, Mansfield and Eliot from Project Gutenberg"
author: Quang Mai
thanks: "Code and data are available at: https://github.com/ponolite/stream-consciousness-language.git"
date: today
date-format: long
abstract: "This project focuses on understanding the language used by renowned stream of consciousness (SOC) authors James Joyce, Virginia Woolf, Marcel Proust, Katherine Mansfield and T.S Eliot. By conducting word frequency analysis and sentiment analysis on these authors' nine novels, this paper aims to uncover shared linguistic patterns and gain insights into the authors' mental states. This paper attempts to offer insights into themes of self-identity, anxiety, disassociation and existential contemplation within Western society and its literary circle from late 19th to mid-20th century. (add one sentence on main results)"
format: pdf
number-sections: true
bibliography: references.bib
toc: true
---

@@

* **Word Frequency:** This variable captures the repetition of a single word (unigram), two-words combination (bigram) or three-words combination (trigram) throughout a SOC novel text, providing us with a thematic understanding of SOC literature.
* **Sentiment Value:** This variable enables us to analyze how every word is usually perceived emotionally, whether it be a qualitative feeling or if it is conveyed through a numerical value.

Out of two variables used, the first one, 'Word Frequency' usually captured as `n` or `frequency` in datasets, is directly quantified through tokenizing the novel texts using the package `tidytext` and its function `unnest_tokens()` [@rTidytext]. To do this, I first downloaded all nine novel texts from @gutenberg, and leveraged functions such as `unnest_tokens()` from @rTidytext to mine the texts, or separating it into individual words. Finally, I used `count()` to quantify the word frequency.

The second variable used, 'Sentiment Value', is based on three English-based, general-purpose  "word-emotion and word-polarity association lexicons", sourced from Mohammad and Turney's expansive research along with efforts from Finn Årup Nielsen and Bing Liu and collaborators [@rTidytext] [@sentiment]. The three general-purpose lexicon that contributes to my sentiment analysis are [@rTidytext]:

@@

In terms of measuring 'Sentiment Value', all three general-purpose lexicons are compiled through crowd-sourcing and directly surveying the public on how each word is emotionally perceived. A survey sample of how the word 'startle' is compiled within the 'nrc' lexicon is presented below [@sentiment]:
